---
title: "Esa's Exploratory Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# Loading & Setting Up Data

```{r}
data <- read_csv("Mental_Health_Care_in_the_Last_4_Weeks.csv")
```

```{r}
glimpse(data)
```

# Beginning exploratory data analysis

The first excursion into this data, yay!

## What does the data look like??

To be honest, the formatting of this data somewhat confuses me and I feel as though I don't understand exactly what all of the variables and conditions are -- this will help fix that!

**Columns**

 - Indicator
 - Group
 - State
 - Subgroup
 - Phase
 - Time Period
 - Time Period Label
 - Time Period Start Date
 - Time Period End Date
 - Value
 - LowCI
 - HighCI
 - Confidence Interval
 - Quartile Range 
 - Suppression Flag *(mostly NA values)*


### What are the possible values for each column?

```{r, eval = FALSE}
#Used to find all of the following values (not actually relevant to any data analysis)
data %>% 
  #filter(`Time Period` == 13, Group == "By State") %>% 
  select("Time Period Start Date") %>% 
  distinct()
```

**Indicators**

The four indicators are:

 - Took Prescription Medication for Mental Health, Last 4 Weeks
 - Received Counseling or Therapy, Last 4 Weeks
 - Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks
 - Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks


**Groups**

The seven groups are:

 - National Estimate	
 - By Age	
 - By Sex	
 - By Presence of Symptoms of Anxiety/Depression	
 - By Race/Hispanic ethnicity	
 - By Education	
 - By State


**States**

The fifty-two 'states' are the entire US population ("United States"), Washington DC ("District of Columbia"), and the fifty US states.


**Subgroups**

The subgroups for each group are:

National Estimate	

- United States	

By Age	

 - 18 - 29 years				
 - 30 - 39 years				
 - 40 - 49 years				
 - 50 - 59 years				
 - 60 - 69 years				
 - 70 - 79 years				
 - 80 years and above	

By Sex	

 - Male				
 - Female				

By Presence of Symptoms of Anxiety/Depression	

 - Did not experience symptoms of anxiety/depression in the past 4 weeks	
  - Experienced symptoms of anxiety/depression in past 4 weeks			

By Race/Hispanic ethnicity	

 - Hispanic or Latino				
 - Non-Hispanic white, single race				
 - Non-Hispanic black, single race				
 - Non-Hispanic Asian, single race				
 - Non-Hispanic, other races and multiple races	

By Education	

 - Less than a high school diploma				
 - High school diploma or GED				
 - Some college/Associate's degree				
 - Bachelor's degree or higher

By State

 - District of Columbia
 - the 50 US states


**Phases**

The three (??) phases are:

 - 2				
 - *NA*			
 - -1
 
 
 **Time Periods**
 
 The sixteen time periods are:

| Period | Period Label     | Period Start Time               |
|-------:|:----------------:|:--------------------------------|
|     13 | 	Aug 19 - Aug 31 | 8/19/20 0:00                    |
|     14 | 	Sep 2 - Sep 14  | 9/2/20 0:00	                    |
|     15 | 	Sep 16 - Sep 28 | 9/16/20 0:00	                  |
|     16 | 	Sep 30 - Oct 12 | 9/30/20 0:00                    |
|     17 | 	Oct 14 - Oct 26 | 10/14/20 0:00			              |
|     18 | 	Oct 28 - Nov 9  | 10/28/20 0:00			              |
|     19 | 	Nov 11 - Nov 23 | 11/11/20 0:00			              |
|     20 | 	Nov 25 - Dec 7  | 11/25/20 0:00			              |
|     21 | 	Dec 9 - Dec 21  | 12/9/20 0:00 OR 12:00:00 AM     |
|     1  |  Dec 22 - Jan 5  | 12/22/20 0:00 OR 12:00:00 AM    |
|     22 |  Jan 6 - Jan 18  | 1/6/21 0:00 OR 12:00:00 AM      |
|     23 |	Jan 20 - Feb 1  | 01/20/2021 12:00:00 AM          |
|     24 |	Feb 3 - Feb 15  | 02/03/2021 12:00:00 AM          |
|     25 |	Feb 17 - Mar 1  | 02/17/2021 12:00:00 AM          |
|     26 |	Mar 3 - Mar 15  | 03/03/2021 12:00:00 AM          |
|     27 |	Mar 17 - Mar 29 | 03/17/2021 12:00:00 AM          |


## Exploring with the first time period, 13 (Aug 19 - Aug 31)

### Map of the US

For this I'm using the packages `usmap`, tutorial from https://cran.r-project.org/web/packages/usmap/vignettes/mapping.html

```{r}
library(ggplot2)
library(usmap)
library(RColorBrewer)
```

```{r, eval=FALSE}
#Reminding myself how to use usmap, mostly taken from tutorial
plot_usmap(regions = "states") + 
  labs(title = "US States",
       subtitle = "This is a blank map of the states of the US (mostly taken from tutorial)") + 
  theme(panel.background = element_rect(color = "black"))
```

```{r, eval=FALSE}
#Also just reminding myself how to use usmap package -- mostly taken from tutorial
plot_usmap(data = statepop, values = "pop_2015", color = "red") + 
  scale_fill_continuous(name = "Population (2015)", label = scales::comma) + 
  theme(legend.position = "right")
```

Let's say we want to look at the indicator "Took Prescription Medication for Mental Health, Last 4 Weeks" for each US state (and DC) in the first time period, 13

```{r}
#Help getting the fips code from https://www.littlemissdata.com/blog/usmap
#Turns out the usmap package won't work without a "state" or "fips" column... and I thought there was a state column but it wasn't working, so fips did the trick!
data_states <- data
data_states$fips <- fips(data_states$State)
```

-1575318.95
-872719.7
```{r}
#Getting the data I want for this
data_states_13_meds <- data_states %>% 
  filter(Indicator == "Took Prescription Medication for Mental Health, Last 4 Weeks",
         Group == "By State",
         `Time Period` == 13) %>% 
  select(fips, Value, State, Indicator) 
```
```{r}
#GETTING ALL OF THE CUSTOM VALUES

#Title and the total value for the united states
full_title <- data_states_13_meds$Indicator %>% head(1)
title <- str_sub(full_title, 1, -15)

national_value <- data_states %>% 
  filter(Indicator == "Took Prescription Medication for Mental Health, Last 4 Weeks",
         Group == "National Estimate",
         `Time Period` == 13) %>% 
  select(Value) %>% 
  head(1) %>% 
  as.numeric()


#Finding the correct color for the total national value
data_states_13_meds$color_lev <- as.numeric(cut(data_states_13_meds$Value, 9, ordered=TRUE))

num <- as.numeric(data_states_13_meds %>% 
  filter(Value == which.min(abs(data_states_13_meds$color_lev-national_value))) %>% select(color_lev))
national_color <- brewer.pal(9, "Blues")[num+1]
```

```{r}
#Now adding the map
plot_usmap(data=data_states_13_meds, values = "Value", color = "white", labels = FALSE, label_color = "white") +  
  scale_fill_distiller(name = "Percent", palette = "Blues", direction = 1) + 
  theme(legend.position = c(.9, .3)) +
  labs(title = title) +
  annotate("text", x = 3000000, y = -1900000, label = "US Total") +
  annotate("point", x = 2825000, y = -2100000, size = 5, color = national_color) +
  annotate("text", x = 3100000, y = -2100000, label = paste(national_value, "%", sep=""))
```

 - Help with plot legend: https://www.statology.org/ggplot-legend-position/
 - Help getting the coordinates and using annotate on the map (I realized the coordinate system is different, so I had to compare the different coordinates to figure out where I could approximately place the legend): https://cran.r-project.org/web/packages/usmap/vignettes/advanced-mapping.html
 - Help with brewer color plot (to find the specific blues in the brewer "Blues" color palette): http://jenrichmond.rbind.io/post/idhtg-how-to-use-colour-palettes-with-ggplot/
  - https://stackoverflow.com/questions/26287864/picking-individual-colours-from-a-rcolorbrewer-palette-as-a-scale-colour-manual
  - http://adomingues.github.io/2015/09/24/finding-closest-element-to-a-number-in-a-list/
  - for bins: https://stackoverflow.com/questions/12979456/categorize-numeric-variable-into-group-bins-breaks and https://www.geeksforgeeks.org/divide-a-vector-into-ranges-in-r-programming-cut-function/ and https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/cut
 - Help with annotation: https://ggplot2.tidyverse.org/reference/annotate.html


Just a map mostly copied from the example, for help finding the correct coordinates:
```{r}
#earthquakes
eq_transformed <- usmap_transform(earthquakes)

eq_transformed <- eq_transformed %>% 
  mutate(lonlat = paste(lon.1, ", ", lat.1))

plot_usmap() +
  geom_point(data = eq_transformed, aes(x = lon.1, y = lat.1, size = mag),
             color = "red", alpha = 0.25) +
  labs(title = "US Earthquakes",
       subtitle = "Source: USGS, Jan 1 to Jun 30 2019",
       size = "Magnitude") +
  geom_label(data = eq_transformed, aes(x = lon.1, y = lat.1, label = lonlat)) +
  theme(legend.position = "right")
```


