---
title: "Final Scatterplot"
subtitle: "SDS 235 Final Project"
author: "Esa Schenck"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    code_folding: hide
    df_print: paged
editor_options: 
  chunk_output_type: inline
---


```{r setup, message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
```

# Set Up

```{r}
#Loading packages
library(tidyverse)
library(RColorBrewer)
library(gridExtra)
library(ggplot2)
```

```{r}
#Loading dataset
new_data_fixed <- read_csv("new_data_fixed.csv")
```


```{r}
#Function to make each plot
make_plot <- function(ind, grp){
  
  caption <- ""
  
  #Indicator things
  title <- ind %>% str_sub(1, -15)
  
  if (ind == "Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks") {
    title <- "Took Medication And/Or Received Counseling or Therapy"
  }
  if (ind == "Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks") {
    title <- "Needed But Didn't Get Counseling"
  }
  
  y_axis_title <- paste("% of Adults Who", title)
  
  #Group things
  if (grp == "By Age") {
    leg_title <- "Age (years)"
    leg_breaks <- c("18 - 29 years", "30 - 39 years", "40 - 49 years", "50 - 59 years", "60 - 69 years", "70 - 79 years", "80 years and above")
    leg_labels <- c("18 - 29", "30 - 39", "40 - 49", "50 - 59", "60 - 69", "70 - 79", "80 +")
    num <- 7
  }
  
  if (grp == "By Sex") {
    leg_title <- "Sex (binary M/F)"
    leg_breaks <- c("Male", "Female")
    leg_labels <- c("Male", "Female")
    caption <- "Note: These labels may not accurately reflect the gender identities of all survey participants"
    num <- 2
  }
  
  if (grp == "By Education") {
    leg_title <- "Education Level"
    leg_breaks <- c("Bachelor's degree or higher", "Some college/Associate's degree", "High school diploma or GED", "Less than a high school diploma")
    leg_labels <- c("Bachelor's", "Some College", "High School", "Less than HS")
    num <- 4
  }
  
  if (grp == "By Presence of Symptoms of Anxiety/Depression") {
    leg_title <- "Symptoms of Anxiety/Depression"
    leg_breaks <- c("Did not experience symptoms of anxiety/depression in the past 4 weeks", "Experienced symptoms of anxiety/depression in past 4 weeks")
    leg_labels <- c("Experienced Symptoms", "No Symptoms")
    caption <- "Note: These data may not accurately reflect the mental health of all survey participants"
    num <- 2
  }
  
  if (grp == "By Race/Hispanic ethnicity") {
    leg_title <- "Race/Hispanic ethnicity"
    leg_breaks <- c("Hispanic or Latino", "Non-Hispanic white, single race", "Non-Hispanic black, single race", "Non-Hispanic Asian, single race", "Non-Hispanic, other races and multiple races")
    leg_labels <- c("Hispanic or Latino", "Non-Hispanic white", "Non-Hispanic black", "Non-Hispanic Asian", "Non-Hispanic, other")
    caption <- "Note: These labels may not accurately reflect the racial/ethnic background of all survey participants"
    num <- 5
  }
  
  my_colors <- brewer.pal(num, "Set1")
  
  #Making the darn plot
  plot <- new_data_fixed %>% 
    filter(Group == grp,
           Indicator == ind) %>% 
    na.omit() %>% #need to omit in order for the geom_ribbon to work
    ggplot(aes(x = avg_cases, y = (Value/100))) +
    geom_ribbon(aes(ymin = (LowCI/100), ymax = (HighCI/100), fill = Subgroup), alpha = .1, show.legend = FALSE) +
    geom_point(aes(color = Subgroup)) +
    labs(title = title, subtitle = "Given the number of COVID cases during the time period", caption = caption) +
    theme(legend.position = "right") +
    scale_color_manual(values = my_colors, name = leg_title, breaks = leg_breaks, labels = leg_labels) +
    scale_fill_manual(values = my_colors, breaks = leg_breaks, labels = leg_labels) +
    theme_classic() +
    scale_x_continuous("Average Daily COVID-19 Cases in the US", labels = scales::comma_format()) +
    scale_y_continuous(y_axis_title, labels = scales::percent_format(accuracy = 5L)) 
  
  plot
}
```




# Variables

**The user can view four different mental health indicators over five different population groups.**

*The mental health indicators are:* US Adults Who...

 - Took Prescription Medication for Mental Health, Last 4 Weeks
 - Received Counseling or Therapy, Last 4 Weeks
 - Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks
 - Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks

*Grouped...*

 - By Age
 - By Sex
 - By Presence of Symptoms of Anxiety/Depression
 - By Race/Hispanic ethnicity
 - By Education

**Here, we're assuming that the user is interested in all of the indicators for one specific grouping.**

# By Age

```{r}
age1 <- make_plot("Took Prescription Medication for Mental Health, Last 4 Weeks", "By Age")
age2 <- make_plot("Received Counseling or Therapy, Last 4 Weeks", "By Age")
age3 <- make_plot("Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks", "By Age")
age4 <- make_plot("Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks", "By Age")

grid.arrange(age1, age2, age3, age4, ncol=2, nrow =2)
```

# By Sex

```{r}
sex1 <- make_plot("Took Prescription Medication for Mental Health, Last 4 Weeks", "By Sex")
sex2 <- make_plot("Received Counseling or Therapy, Last 4 Weeks", "By Sex")
sex3 <- make_plot("Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks", "By Sex")
sex4 <- make_plot("Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks", "By Sex")
```
```{r}
grid.arrange(sex1, sex2, sex3, sex4, ncol=2, nrow =2)
```

# By Presence of Symptoms of Anxiety/Depression
 
```{r}
symptoms1 <- make_plot("Took Prescription Medication for Mental Health, Last 4 Weeks", "By Presence of Symptoms of Anxiety/Depression")
symptoms2 <- make_plot("Received Counseling or Therapy, Last 4 Weeks", "By Presence of Symptoms of Anxiety/Depression")
symptoms3 <- make_plot("Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks", "By Presence of Symptoms of Anxiety/Depression")
symptoms4 <- make_plot("Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks", "By Presence of Symptoms of Anxiety/Depression")
```
```{r}
grid.arrange(symptoms1, symptoms2, symptoms3, symptoms4, ncol=2, nrow =2)
```

# By Race/Hispanic ethnicity

```{r}
race1 <- make_plot("Took Prescription Medication for Mental Health, Last 4 Weeks", "By Race/Hispanic ethnicity")
race2 <- make_plot("Received Counseling or Therapy, Last 4 Weeks", "By Race/Hispanic ethnicity")
race3 <- make_plot("Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks", "By Race/Hispanic ethnicity")
race4 <- make_plot("Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks", "By Race/Hispanic ethnicity")
```
```{r}
grid.arrange(race1, race2, race3, race4, ncol=2, nrow =2)
```

# By Education

```{r}
education1 <- make_plot("Took Prescription Medication for Mental Health, Last 4 Weeks", "By Education")
education2 <- make_plot("Received Counseling or Therapy, Last 4 Weeks", "By Education")
education3 <- make_plot("Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks", "By Education")
education4 <- make_plot("Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks", "By Education")
```
```{r}
grid.arrange(education1, education2, education3, education4, ncol=2, nrow =2)
```

