---
title: "Explore"
author: "Nage Ngo"
date: "5/9/2021"
output:   
  html_document:
    toc: true
    toc_depth: 3
    theme: cerulean
    highlight: tango
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, error = F, message = F)
library(plotly)
library(tidyverse)
library(ggplot2)
# library(usmap)
# library(RColorBrewer)
library(maps)     # Provides latitude and longitude data for various maps
```

```{r import}
data <- read_csv("Mental_Health_copy.csv", 
                 col_types = cols(`Quartile Range` = col_skip(), 
                                  `Suppression Flag` = col_skip(),
                                  LowCI = col_skip(),
                                  HighCI= col_skip(),
                                  `Confidence Interval`= col_skip())) %>%
  mutate(t = ifelse(`Time Period`>21, `Time Period`+1,`Time Period`)) %>%
  mutate(t2 = ifelse(t==1, 22,t)) %>%
  select(-c(t,`Time Period`)) %>% rename(`Time Period`=t2)
```

# Temporal by state and indicator

### Took Prescription Medication for Mental Health, Last 4 Weeks

```{r}
# taking prescription filter
data_states_13_meds <- data %>% 
  filter(Indicator == "Took Prescription Medication for Mental Health, Last 4 Weeks",
         Group == "By State" ) %>% 
  mutate(State1 = tolower(State)) %>%
  select(Value, State = State1, Indicator, t = `Time Period`, `Time Period Label`) 

```


```{r}
MainStates <- map_data("state") %>% mutate(State = tolower(region)) 
merged_pres <- inner_join(MainStates, data_states_13_meds, by = "State")

# MainStates_lab <- MainStates %>% distinct(region,.keep_all=T)
region.lab.data <- merged_pres %>%
  group_by(region) %>% summarise(long=mean(long), lat=mean(lat))

map_took_pres <- ggplot() + 
  theme_classic() +
  geom_polygon(data=merged_pres, 
          aes(x=long, y=lat, group=group, fill = Value, frame =  t), 
          color="white", size = 0.2) +
  scale_fill_gradient(low = "lightblue", high = "darkred")  +
  geom_text(aes(label = region,x = long, y = lat), data = region.lab.data, size=1, alpha = 0.001) +
  # annotate("text", x = -120, y = 25, label = merged_pres$`Time Period Label`, frame = t) +
  labs(title="Took Prescription Medication for Mental Health, Last 4 Weeks", fill = "%")+
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank())

fig <- ggplotly(map_took_pres) 
fig <- fig %>% 
  animation_opts(
    0, easing = "elastic", redraw = FALSE
  ) %>%
  animation_slider(
    currentvalue = list(prefix = "Time Period ", font = list(color="red"))
  ) #%>%
  # layout(annotations = list(x = -120, y = 25, text = paste("Time Period: ",
  #   merged_pres$`Time Period Label`), showarrow = F))
  

fig$x$frames <- lapply(
  fig$x$frames, function(f) { 
    f$data <- lapply(f$data, function(d) d[!names(d) %in% c("x", "y")])
    f 
  })

fig
```

### Received Counseling or Therapy, Last 4 Weeks

```{r}
data_states_13_thrp <- data %>% 
  filter(Indicator == "Received Counseling or Therapy, Last 4 Weeks",
         Group == "By State") %>% 
  mutate(State1 = tolower(State)) %>%
  select(Value, State = State1, Indicator, t = `Time Period`, `Time Period Label`) 

# merge with state
merged_thrp <- inner_join(MainStates, data_states_13_thrp, by = "State")

map_thrp <- ggplot() + 
  theme_classic() +
  geom_polygon(data=merged_thrp, 
          aes(x=long, y=lat, group=group, fill = Value, frame =  t), 
          color="white", size = 0.2) +
  scale_fill_gradient(low = "lightblue", high = "darkred")  +
  geom_text(aes(label = region,x = long, y = lat), data = region.lab.data, size=1, alpha = 0.001) +
  labs(title="Received Counseling or Therapy, Last 4 Weeks", fill = "%")+
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank())

fig2 <- ggplotly(map_thrp) 
fig2 <- fig2 %>% 
  animation_opts(
    0, easing = "elastic", redraw = FALSE
  )

fig2$x$frames <- lapply(
  fig2$x$frames, function(f) { 
    f$data <- lapply(f$data, function(d) d[!names(d) %in% c("x", "y")])
    f 
  })

fig2
```

### Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks

```{r}
data_states_13_both <- data %>% 
  filter(Indicator == "Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks",
         Group == "By State" 
         #,`Time Period` == 13
         ) %>% 
  mutate(State1 = tolower(State)) %>%
  select(Value, State = State1, Indicator, t = `Time Period`, `Time Period Label`) 

# merge with state
merged_both <- inner_join(MainStates, data_states_13_both, by = "State")

map_both <- ggplot() + 
  theme_classic() +
  geom_polygon(data=merged_both, 
          aes(x=long, y=lat, group=group, fill = Value, frame =  t), 
          color="white", size = 0.2) +
  scale_fill_gradient(low = "lightblue", high = "darkred")  +
  # geom_text(data=MainStates_lab, aes(long, lat, label = region), size=2) +
  labs(title="Took Prescription Medication for Mental Health And/Or Received Counseling or Therapy, Last 4 Weeks", fill = "%")+
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank())

fig3 <- ggplotly(map_both) 
fig3 <- fig3 %>% 
  animation_opts(
    0, easing = "elastic", redraw = FALSE
  )

fig3$x$frames <- lapply(
  fig3$x$frames, function(f) { 
    f$data <- lapply(f$data, function(d) d[!names(d) %in% c("x", "y")])
    f 
  })

fig3
```

### Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks

```{r}
data_states_13_need <- data %>% 
  filter(Indicator == "Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks",
         Group == "By State" 
         #,`Time Period` == 13
         ) %>% 
  mutate(State1 = tolower(State)) %>%
  select(Value, State = State1, Indicator, t = `Time Period`, `Time Period Label`) 

# merge with state
merged_need <- inner_join(MainStates, data_states_13_need, by = "State")

map_need <- ggplot() + 
  theme_classic() +
  geom_polygon(data=merged_need, 
          aes(x=long, y=lat, group=group, fill = Value, frame =  t), 
          color="white", size = 0.2) +
  scale_fill_gradient(low = "lightblue", high = "darkred")  +
  # geom_text(data=MainStates_lab, aes(long, lat, label = region), size=2) +
  labs(title="Needed Counseling or Therapy But Did Not Get It, Last 4 Weeks", fill = "%")+
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank())

fig4 <- ggplotly(map_need) 
fig4 <- fig4 %>% 
  animation_opts(
    0, easing = "elastic", redraw = FALSE
  )

fig4$x$frames <- lapply(
  fig4$x$frames, function(f) { 
    f$data <- lapply(f$data, function(d) d[!names(d) %in% c("x", "y")])
    f 
  })

fig4
```



# Period mismatch

```{r}
data%>% 
  distinct(`Time Period`, 
           `Time Period Label`)
```
Mar 17 - Mar 29	 --> 27
Dec 22 - Jan 5	--> 1 --> between 21 and 22


**period numbers mismatch**

**add national averages**

**add state label**








# Hover for drop down

By indicator, Drop down info: state, gender, race

```{r, eval = F}

data_state_pres <- data %>% 
  filter(Indicator == "Took Prescription Medication for Mental Health, Last 4 Weeks")%>%
  pivot_wider(names_from = c("Subgroup"), values_from = "Value") %>%
  distinct(.keep_all = T)
# data$hover <- with(data, paste(State, '<br>', "Female", Female, "Male", Male, "<br>",
#                                "Hispanic or Latino", `Hispanic or Latino`, "<br>",
#                                "White", `Non-Hispanic white, single race`,
#                                "Asian", `Non-Hispanic Asian, single race`,"<br>",
#                                "Black", `Non-Hispanic black, single race`, "Other/Multiple Races", 
#                                `Non-Hispanic, other races and multiple races`))
# 
# fig <- plot_geo(df, locationmode = 'USA-states')
# fig <- fig %>% add_trace(
#     z = ~total.exports, text = ~hover, locations = ~code,
#     color = ~total.exports, colors = 'Purples'
#   )
# fig <- fig %>% colorbar(title = "Millions USD")
# fig <- fig %>% layout(
#     title = '2011 US Agriculture Exports by State<br>(Hover for breakdown)'
#   )
# 
# fig
```


